---
title: "CFF_phylo"
author: "Kevin Healy"
date: "21/05/2021"
output: html_document
---




```{r packages, tidy = TRUE, results = "hide"}

require(mulTree)

library(devtools)
library(MASS)
library(phytools)
library(caper)
library(taxize)
library(fishtree)
library(rotl)
library(phangorn)
```

Lets also load some functions that we built specifically for this analysis 
The phylo_bind_functions script has functions to build the phylogenies and include error in the nodes joining together trees.

```{r load additional functions, include=TRUE}

source("phylo_bind_functions.R")

```



Read in the data

```{r setup, include=FALSE}

cff_raw <- read.csv("Cff_main_data_pre_phylo_22_July_2021.csv", 
                    header = T,
                    sep = ",")

 cff_raw[cff_raw$mode_of_life == "parasite",]$mode_of_life <- "pred_forage"

 cff_raw[cff_raw$mode_of_life == "scavenger",]$mode_of_life <- "pred_forage"

 cff_raw[cff_raw$mode_of_life == "herbivore",]$mode_of_life <- "pred_forage"


 cff_raw[cff_raw$Habitat == "marine_reef",]$Habitat <- "demersal"

 cff_raw[cff_raw$Habitat == "bathypelagic",]$Habitat <- "pelagic"



cff_raw$mode_of_life <- factor(cff_raw$mode_of_life, 
                                  levels = c("pred_forage",   
                                             "deposite", 
                                             "pred_ballistic",   
                                            #  "scavenger",  
                                             "pred_pursuit"
                                            #  ,"herbivore"
                                            # ,"parasite"
                                             ))


cff_raw$Habitat <- factor(cff_raw$Habitat, 
                                  levels = c("terrestrial",   
                                             "pelagic", 
                                             "demersal",   
                                             #"marine_reef",
                                             #"bathypelagic",
                                             "volant",
                                             "arboreal",
                                             "benthopelagic"
                                             ))
  
 
cff_clean_no_syn <- cff_raw[,c("Scientific_name",
                        "taxa_group",
                        "taxa_phyla",
                        "taxa_class",
                        "Light_Level",
                        "Method",
                        "Max_CFF_Hz",
                        "Bodymass.g.",
                        "Habitat",
                        "mode_of_life")]
 
cff_clean_no_syn <- na.omit(cff_clean_no_syn)

```



#Rename some species names to match the tree

```{r setup, include=FALSE}


cff_clean <- cff_clean_no_syn

#cff_clean <- cff_clean[cff_clean$Scientific_name != "Tamias_ameonus", ] 


#Glossina_morsitans to XX
cff_clean[cff_clean$Scientific_name == "Glossina_morsitans", "Scientific_name"] <- "Glossina_morsitans_morsitans"

#Cyanistes_caeruleus to XX
cff_clean[cff_clean$Scientific_name == "Cyanistes_caeruleus", "Scientific_name"] <- "Parus_caeruleus"


#Acromegalomma_vesiculosum to Branchiomma_vesiculosum
cff_clean[cff_clean$Scientific_name == "Acromegalomma_vesiculosum", "Scientific_name"] <- "Branchiomma_vesiculosum"

#Anisoptera, which is an infroorder of dragonflies to be represented by Sympetrum_flaveolum_ott906647
cff_clean[cff_clean$Scientific_name == "Anisoptera", "Scientific_name"] <- "Sympetrum_flaveolum"


#anolis_carolinensis to Anolis_carolinensis
cff_clean[cff_clean$Scientific_name == "anolis_carolinensis", "Scientific_name"] <- "Anolis_carolinensis"


#anolis_limifrons to Anolis_limifrons
cff_clean[cff_clean$Scientific_name == "anolis_limifrons", "Scientific_name"] <- "Anolis_limifrons"


#anolis_sagrei to Anolis_sagrei
cff_clean[cff_clean$Scientific_name == "anolis_sagrei", "Scientific_name"] <- "Anolis_sagrei"


#anolis_valencienni to Anolis_valencienni
cff_clean[cff_clean$Scientific_name == "anolis_valencienni", "Scientific_name"] <- "Anolis_valencienni"


#Cyanistes_caeruleus to Parus_caeruleus
cff_clean[cff_clean$Scientific_name == "cyanistes_caeruleus", "Scientific_name"] <- "Parus_caeruleus"


#Caretta_caretta to Caretta_caretta_HBS112700
cff_clean[cff_clean$Scientific_name == "Caretta_caretta", "Scientific_name"] <- "Caretta_caretta_HBS112700"


#Chelonia_mydas to Chelonia_mydas_HBS109887
cff_clean[cff_clean$Scientific_name == "Chelonia_mydas", "Scientific_name"] <- "Chelonia_mydas_HBS109887"

#Dermochelys_coriacea to Dermochelys_coriacea_MVZ149847
cff_clean[cff_clean$Scientific_name == "Dermochelys_coriacea", "Scientific_name"] <- "Dermochelys_coriacea_MVZ149847"


#Felis_domesticus to Felis_silvestris
cff_clean[cff_clean$Scientific_name == "Felis_domesticus", "Scientific_name"] <- "Felis_silvestris"


#Gekko_gekko to Gekko_gecko
cff_clean[cff_clean$Scientific_name == "Gekko_gekko", "Scientific_name"] <- "Gekko_gecko"


#Janicella_spinacauda to Janicella_spinicauda
cff_clean[cff_clean$Scientific_name == "Janicella_spinacauda", "Scientific_name"] <- "Janicella_spinicauda"


#Lycosa_baltimoriana to Lycosa_beihaiensis
cff_clean[cff_clean$Scientific_name == "Lycosa_baltimoriana", "Scientific_name"] <- "Lycosa_beihaiensis"


#Nematobrachian_boopis to Nematobrachion_boopis
cff_clean[cff_clean$Scientific_name == "Nematobrachian_boopis", "Scientific_name"] <- "Nematobrachion_boopis"

#Nematobrachion_sexpinosus to Nematobrachion_sexspinosum
cff_clean[cff_clean$Scientific_name == "Nematobrachion_sexpinosus", "Scientific_name"] <- "Nematobrachion_sexspinosum"


#Apis_mellifera, to Apis_mellifera_mellifera_ott13206
cff_clean[cff_clean$Scientific_name == "Apis_mellifera", "Scientific_name"] <- "Apis_mellifera_mellifera"

#Axcanthaster_planci, to Acanthaster_planci_ott926817
cff_clean[cff_clean$Scientific_name == "Axcanthaster_planci", "Scientific_name"] <- "Acanthaster_planci"

#Periplaneta_americano, to Periplaneta_americana
cff_clean[cff_clean$Scientific_name == "Periplaneta_americano", "Scientific_name"] <- "Periplaneta_americana"


#salmo_salar, to Salmo_salar
cff_clean[cff_clean$Scientific_name == "salmo_salar", "Scientific_name"] <- "Salmo_salar"

#Rynchophorus_ferrugineus, to Rhynchophorus_ferrugineus
cff_clean[cff_clean$Scientific_name == "Rynchophorus_ferrugineus", "Scientific_name"] <- "Rhynchophorus_ferrugineus"


#xiphias_gladius, to Xiphias_gladius
cff_clean[cff_clean$Scientific_name == "xiphias_gladius", "Scientific_name"] <- "Xiphias_gladius"


#Calliphoridae, which is all blowflies to be represented by Chrysomya_megacephala
cff_clean[cff_clean$Scientific_name == "Calliphoridae", "Scientific_name"] <- "Chrysomya_megacephala"


#Change Canis_lupus_familiaris to Canis_lupus
cff_clean[cff_clean$Scientific_name == "Canis_lupus_familiaris", "Scientific_name"] <- "Canis_lupus"


#Change Gallus_gallus_domesticus to Gallus_gallus
cff_clean[cff_clean$Scientific_name == "Gallus_gallus_domesticus", "Scientific_name"] <- "Gallus_gallus"


#Change Iguana_Iguana to Iguana_iguana
cff_clean[cff_clean$Scientific_name == "Iguana_Iguana", "Scientific_name"] <- "Iguana_iguana"


#Change locusta_migratoria to Locustaevora_migratoriae
cff_clean[cff_clean$Scientific_name == "locusta_migratoria", "Scientific_name"] <- "Locustaevora_migratoriae"

#Change Raja_erinacea to Leucoraja_erinacea
#cff_clean[cff_clean$Scientific_name == "Raja_erinacea", "Scientific_name"] <- "Leucoraja_erinacea"


```

Lets now upload each of the phylogenies we are going to use to build our super-dooper tree.
We used the Metazoan phylogeny from the Open Tree of Life project (https://ot39.opentreeoflife.org/opentree/argus/opentree10.4@ott93302)  as a backbone. For Mammals we used a distribution of 100 trees from Kuhn et al 2011. For Aves we used the Jetz et al 2012 distribution of trees. For lizards we used the Pyron et al 2014 phylogeny.

For our analysis we will use a distribution of 100 supertrees, however as many of the trees will not be suitable (for example, when first apperance fossil ranges overlap between nodes resulting in negative distances for certain topologies.)

*Please note that as the seperate phylogenies are too large to host on Github they will need to be downloaded seperately from there sources. However, we do include the output distribution of trees (COMADRE_100_phylo_feb_2019.tre) that is the result of this script within the repository.


```{r load trees, include=TRUE, warning = FALSE}

setwd("/Users/kh214/Desktop/Phlyogenies")

#Metazoan backbone
met_tree <- read.tree("metazoa.tre")
met_tree  <- makeLabel(met_tree) 
met_tree <- drop.tip(met_tree,"Homo_sapiens")

met_tree$node.label<-NULL

#Mammal phylogeny
mammal_phy <- read.nexus("FritzTree.rs200k.100trees.tre")

#Aves phylogeny
aves_phy <- read.tree("BirdzillaHackett10.tre")

#Lizard phylogeny
Lizard_phy <- read.tree("liz_and_snake_time_tree.txt")
lizard_phy<-makeLabel(Lizard_phy)
#this removes the node lables which can cause dublicate problems
lizard_phy$node.label<-NULL
#this fixes rounding errors due to format change to make it ultrametric.
lizard_phy <-chronoMPL(lizard_phy)


#Turtle phylogeny

Turt_phy <- read.nexus("Turtle_Posterior_100.tre")


#amphiban phylo
amph_phy <- read.tree("Amphi.tre")

amph_phy <- drop.tip(amph_phy, "Homo_sapiens")

#this removes the node lables which can cause dublicate problems
amph_phy$node.label<-NULL
amph_phy <-chronoMPL(amph_phy)


#Shark
shark_phy <- read.nexus("Shark_10.cal.tree.nex")


# lets also set the number of trees we will build. 
# As we will have to drop trees due to the inclusion of uncertinatly in node age 
# creating trees that dont corresponse with the structure of our backbone phylogeny
# we need to more than 100
sample.no <- 1000


setwd("/Users/kh214/Desktop/GitHub/CFF_full_animal")
```


Now lets creat some super trees. First we join birds and turtles together


For the common ancestor for turtles and aves we use the stem dates proposed by W. G. Joyce, J. F. Parham, T. R. Lyson, R. C. M. Warnock, P. C. J. Donoghue, A divergence dating analysis of turtles using fossil calibrations: An example of best practices.J. Paleontol. 87, 612–634 (2013).


```{r join aves and squamata, include=TRUE, warning = FALSE}

diapsid_o <- jiggle.bind(x = aves_phy, 
                               y = Turt_phy, 
                               sample = 1000, 
                               min.age = 156, 
                               max.age = 250)


diapsid_nultra <- diapsid_o[[1]]
diapsid_node <- diapsid_o[[2]]


is_ultra_diapsid <- vector()
for(i in 1:length(diapsid_nultra)) {
  is_ultra_diapsid[i] <- is.ultrametric(diapsid_nultra[[i]])
  }

ultra_row_d <- which(is_ultra_diapsid ==TRUE)
diapsid <- diapsid_nultra[ultra_row_d]

class(diapsid) <- "multiPhylo"



```





To link these trees to the common ancestor of Lepidosauria we used a range of 259-285 Myr from Jones et al 2013 

```{r join aves and squamata, include=TRUE, warning = FALSE}

length_bird_lizard <- as.numeric(length(diapsid))

bird_lizard_o <- jiggle.bind(x = diapsid, 
                               y = lizard_phy, 
                               sample = length_bird_lizard, 
                               min.age = 259, 
                               max.age = 285)


bird_lizard_nultra <- bird_lizard_o[[1]]
bird_lizard_node <- bird_lizard_o[[2]]


is_ultra_bird_lizard<- vector()
for(i in 1:length(bird_lizard_nultra)) {
  is_ultra_bird_lizard[i] <- is.ultrametric(bird_lizard_nultra[[i]])
  }

ultra_row_bl <- which(is_ultra_bird_lizard ==TRUE)
bird_lizard <- bird_lizard_nultra[ultra_row_bl]

class(bird_lizard) <- "multiPhylo"

```




For the common ancestor of amniotes, we used the fossil Archerpeton anthracos (Holotype: RM 12056, Author: Carroll 1964, Reisz and Müller, Epoch: Westphalian A Canada Nova Scotia, Age: 318.1 – 314.6 Myr, Dating: International Commission on Stratigraphy 2009)

```{r join aves and squamata to mammals, include=TRUE, warning = FALSE}

length_amniote <- as.numeric(length(bird_lizard))

amniote_tree_o <- jiggle.bind(x = bird_lizard, 
                             y = mammal_phy, 
                             sample = length_amniote, 
                             min.age = 314.6, 
                             max.age = 318.1)


amniote_tree_nultra <- amniote_tree_o[[1]]
amniote_node <- amniote_tree_o[[2]]


is_ultra_amniote_tree <- vector()
for(i in 1:length(amniote_tree_nultra)) {
  is_ultra_amniote_tree[i] <- is.ultrametric(amniote_tree_nultra[[i]])
  }

ultra_rowam <- which(is_ultra_amniote_tree ==TRUE)
amniote_tree <- amniote_tree_nultra[ultra_rowam]


class(amniote_tree) <- "multiPhylo"

```



```{r join amniotes to amphibians, include=TRUE, warning = FALSE}


length_amni_amphi <- as.numeric(length(amniote_tree))


amni_amphi_tree_o <- jiggle.bind(x = amniote_tree, 
                             y = amph_phy, 
                             sample = length_amni_amphi, 
                             min.age = 359.2, 
                             max.age = 359.2)


amni_amphi_tree_nultra <- amni_amphi_tree_o[[1]]
amni_amphi_node <- amni_amphi_tree_o[[2]]


is_ultra_amni_amphi_tree <- vector()
for(i in 1:length(amni_amphi_tree_nultra)) {
  is_ultra_amni_amphi_tree[i] <- is.ultrametric(amni_amphi_tree_nultra[[i]])
  }

ultra_rowaa <- which(is_ultra_amni_amphi_tree ==TRUE)
amni_amphi_tree <- amni_amphi_tree_nultra[ultra_rowaa]


class(amni_amphi_tree) <- "multiPhylo"

```



Next we add fish


```{r join tetrapods to fish, include=TRUE, warning = FALSE}

fish_data <- cff_clean[cff_clean$taxa_phyla %in% c("fish"),]
fish_data <- data.frame(fish_data, species_match = fish_data$Scientific_name )


w_fish_tree <- fishtree_phylogeny(type = c("chronogram"))


fish_tree <- comparative.data(phy = w_fish_tree, 
                              data = fish_data, 
                              names.col = "species_match" , 
                              force.root = TRUE)$phy


##Need to get a sensable split. age should be at lungfish or somerthing
#tree_fish $edge.length <- tree_fish $edge.length*418.5

length_fish <- as.numeric(length(amni_amphi_tree))


fish_tetrapod_tree_o <- jiggle.bind(x = amni_amphi_tree, 
                                    y = fish_tree, 
                                    sample = length_fish, 
                                    min.age = 419, 
                                    max.age = 419)


fish_tetrapod_tree_nultra <- fish_tetrapod_tree_o[[1]]
fish_tetrapod_node <- fish_tetrapod_tree_o[[2]]


is_ultra_fish_tetrapod_tree <- vector()
for(i in 1:length(fish_tetrapod_tree_nultra)) {
  is_ultra_fish_tetrapod_tree[i] <- is.ultrametric(fish_tetrapod_tree_nultra[[i]])
  
  fish_tetrapod_tree_nultra[[i]]$node.label = "NA"

  }

ultra_rowff <- which(is_ultra_fish_tetrapod_tree ==TRUE)
fish_tetrapod_tree <- fish_tetrapod_tree_nultra[ultra_rowff]


class(fish_tetrapod_tree) <- "multiPhylo"

```

Next we add sharks


```{r join tetrapods to fish, include=TRUE, warning = FALSE}

shark_data <- cff_clean[cff_clean$taxa_phyla %in% c("elasmo"),]
shark_data <- data.frame(shark_data, species_match = shark_data$Scientific_name )


length_shark <- as.numeric(length(fish_tetrapod_tree))


shark_tree <- list()
for(i in 1:length_shark){
shark_tree[[i]] <- comparative.data(phy = shark_phy[[i]], 
                              data = shark_data, 
                              names.col = "species_match" , 
                              force.root = TRUE)$phy

}

class(shark_tree) <- "multiPhylo"

##Need to get a sensable split. for sharks

shark_fish_tree_o <- jiggle.bind(x = fish_tetrapod_tree, 
                                    y = shark_tree, 
                                    sample = length_shark, 
                                    min.age = 420, 
                                    max.age = 420)


shark_fish_tree_nultra <- shark_fish_tree_o[[1]]
shark_fish_tree_node <- shark_fish_tree_o[[2]]

is_ultra_shark_fish_tree <- vector()
for(i in 1:length(shark_fish_tree_nultra)) {
  is_ultra_shark_fish_tree[i] <- is.ultrametric(shark_fish_tree_nultra[[i]])
  
  shark_fish_tree_nultra[[i]]$node.label = "NA"
  
  }

ultra_rowss <- which(is_ultra_shark_fish_tree ==TRUE)
shark_fish_tree <- shark_fish_tree_nultra[ultra_rowss]


class(shark_fish_tree) <- "multiPhylo"

```


Add in the river lampray. used time tree for divergance using Cyclostomata 434 ± 8 from Delsuc et al 2018 A phylogenomic framework and timescale
for comparative studies of tunicates.

```{r join lampray to vertabrates, include=TRUE, warning = FALSE}




lamp_data <- cff_clean[cff_clean$Scientific_name %in% c("Lampetra_fluviatilis",
                                                      "Mordacia_praecox",
                                                      "Mordacia_mordax"),]


met_tree_l <- met_tree

for(i in 1:(length(lamp_data$Scientific_name))){
  
  met_tree_l$tip.label[grep(lamp_data$Scientific_name[i],met_tree_l$tip.label)] <- sub("_ott.*", "",met_tree$tip.label[grep(lamp_data$Scientific_name[i],met_tree$tip.label)])
  
}


lamp_tree <- comparative.data(phy = met_tree_l, 
                              data = lamp_data, 
                              names.col = "Scientific_name" , 
                              force.root = TRUE)$phy


class(lamp_tree)<-"phylo"
lamp_tree$node.label = "NA"


length_lamp <- as.numeric(length(shark_fish_tree))

chordata_tree_o <- jiggle.bind(x = lamp_tree, 
                                    y = shark_fish_tree, 
                                    sample = length_lamp, 
                                    min.age = 426, 
                                    max.age = 442)

chordata_tree <- chordata_tree_o[[1]]

for( i in 1:length_lamp){
chordata_tree[[i]]$node.label = "NA"
}

chordata_tree_node <- chordata_tree_o[[2]]
class(chordata_tree) <- "multiPhylo"

```

No add in the inverr groups.

First add the echinoderm.  Ambulacraria 551 ± 16 Delsuc et al 2018 A phylogenomic framework and timescale
for comparative studies of tunicates.

```{r join lampray to vertabrates, include=TRUE, warning = FALSE}


echino_tip <-list(edge=matrix(c(2,1),1,2),
           tip.label= "Acanthaster_planci",
           edge.length=1.0,
           Nnode=1)

class(echino_tip)<-"phylo"

length_echino <- as.numeric(length(chordata_tree))

echino_tree_o <- jiggle.bind(x = echino_tip, 
                                    y = chordata_tree, 
                                    sample = length_echino, 
                                    min.age = 535, 
                                    max.age = 567)

echino_tree <- echino_tree_o[[1]]

for( i in 1:length_echino){
echino_tree[[i]]$node.label = "NA"
}

echino_tree_node <- echino_tree_o[[2]]
class(echino_tree) <- "multiPhylo"

```



For the common ancestor between deuterostomes and protostomes we 599 ± 11 from Delsuc et al 2018 A phylogenomic framework and timescale
for comparative studies of tunicates.

```{r join verts to inverts, include=TRUE, warning = FALSE}

invert_data <- cff_clean[cff_clean$taxa_group == "invert",]
invert_data <- invert_data[invert_data$Scientific_name != "Acanthaster_planci",]


met_tree2 <- met_tree

for(i in 1:(length(invert_data$Scientific_name))){
  
  met_tree2$tip.label[grep(invert_data$Scientific_name[i],met_tree2$tip.label)] <- sub("_ott.*", "",met_tree$tip.label[grep(invert_data$Scientific_name[i],met_tree$tip.label)])
  
}



invert_tree <- comparative.data(phy = met_tree2, 
                              data = invert_data, 
                              names.col = "Scientific_name" , 
                              force.root = TRUE)

invert_tree_ult <-compute.brlen(invert_tree$phy, 
                          method = "Grafen", 
                          power = 1)

##Need to get a sensable split. age should be at lungfish or somerthing
invert_tree_ult$edge.length <- invert_tree_ult $edge.length*587

length_bi <- as.numeric(length(chordata_tree))

bi_lat_tree_o <- jiggle.bind(x = invert_tree_ult, 
                                    y = echino_tree, 
                                    sample = length_bi, 
                                    min.age = 588, 
                                    max.age = 610)


bi_lat_tree <- bi_lat_tree_o[[1]]


for( i in 1:length_bi){
bi_lat_tree[[i]]$node.label = "NA"
}


bi_lat_node <- bi_lat_tree_o[[2]]
class(bi_lat_tree) <- "multiPhylo"

```




```{r building the final trees, include=TRUE, warning = FALSE}

final_tree<- list()

##Now we clean each of the trees so it only has the species from the dataset in
##it
for(i in 1:(length(bi_lat_tree))){
  
  final_tree[[i]] <- comparative.data(phy = bi_lat_tree[[i]], 
                                      data = cff_clean, 
                                      names.col = "Scientific_name", 
                                      force.root = TRUE)$phy
}


final_cff_data <- comparative.data(phy = bi_lat_tree[[1]], 
                                      data = cff_clean, 
                                      names.col = "Scientific_name", 
                                      force.root = TRUE)


class(final_tree) <- "multiPhylo"


```



Write the data and the tree

```{r write, include=TRUE, warning = FALSE}

write.tree(final_tree, file = "cff_tree_27_july_2021.tree")


write.csv(file = "final_cff_data_27_july_2021.csv", final_cff_data$data)

```
