---
title: "New_cff"
author: "Kevin Healy"
date: "17/11/2020"
output: html_document
---



```{r packages, tidy = TRUE, results = "hide"}

library(phytools)
library(caper)
library(MCMCglmm)
library(hdrcde)
library(mulTree)
#source("MultiDisPlot.R")
```


Read in the data

```{r setup, include=FALSE}

cff_raw <- read.csv("Full_CFF_working_12_7_24.csv", 
                    header = T,
                    sep = ",")

#This drops a couple of studies that are not up to scratch as measures
cff_raw <- cff_raw[cff_raw$Critical_appraisal_keep == "y",]

#Set forage as the baseline
cff_raw$mode_of_life <- factor(cff_raw$mode_of_life, 
                                  levels = c("forage",   
                                             "ballistic",   
                                             "pursuit"
                                             ))

#Set terrestrial as the baseline
cff_raw$habitat <- factor(cff_raw$habitat, 
                                  levels = c("terrestrial",
                                             "pelagic", 
                                             "demersal",
                                             "volant"
                                             ))
  
 
cff_clean <- cff_raw[,c("Species_phylo",
                        "Class",
                        "habitat",
                        "forage_light_level",
                        "body_mass_us",
                        "mode_of_life",
                        "Method",
                        "CFF")]
 

cff_clean <- data.frame(cff_clean, animal = cff_clean$Species_phylo)

cff_clean <- na.omit(cff_clean)

```


Read in the phylogeny.

```{r phylo, include=FALSE}

cff_tree <- read.tree("cff_tree_7_8_2024.tree")

cff_tree <- cff_tree[1:100]

```

We can plot the phylogeny. I need to fix the barplot

```{r phylo plot, include=FALSE}


cff_phylo_plot_cff <- c()
cff_phylo_plot_light <- c()
cff_phylo_plot_species <- c()

for(i in 1:length(unique(cff_clean$Species_phylo))){
  
 cff_phylo_plot_species[i] <-  unique(cff_clean$Species_phylo)[i]
 cff_phylo_plot_light[i] <-  cff_clean[cff_clean$Species_phylo == unique(cff_clean$Species_phylo)[i], "forage_light_level"][1]
 cff_phylo_plot_cff[i] <-  max(cff_clean[cff_clean$Species_phylo == unique(cff_clean$Species_phylo)[i], "CFF"])

}


cff_phylo_plot <- data.frame(cff_phylo_plot_species,
                             cff_phylo_plot_light,
                             cff_phylo_plot_cff)

#create data with rownames for p
rownames(cff_phylo_plot) <- cff_phylo_plot$cff_phylo_plot_species

Max_CFF_Hz_plot <-setNames(cff_phylo_plot$cff_phylo_plot_cff,rownames(cff_phylo_plot))


par(mar=c(4.1,4.1,4.1,2.1))

plotTree.wBars(cff_tree[[1]], 
                 Max_CFF_Hz_plot,
     type = "fan",
     scale = 0.5,
     tip.labels = F, cex = 0.2)


plot(cff_tree[[1]],
          type = "fan",
          cex = 0.5)

```



#MulTree
Set the data into a Multree object so we can run the MCMCglmm over 100 
phylogenies once we are ready

```{r Multree data, include=FALSE}

cff_mul <- as.mulTree(cff_clean, 
                      cff_tree, 
                      taxa = "animal", 
                      rand.terms = ~animal,
                      clean.data = FALSE)


```



##Set the parameters
Here we set the number of iterations, the thinning and the burn-in

```{r parameters}

nitt <- 240000
thin <- 100
burnin <- 40000

parameters <- c(nitt,
                thin,
                burnin)

```


##Set the prior

Set the prior.

```{r prior}
prior <-list(R = list(V = 1, nu=0.002), 
             G = list(G1=list(V = 1,n = 2, 
                              nu=0.002),
                      G2=list(V = 1,n = 2, 
                              nu=0.002)
                     ))


```


```{r glm, include=FALSE}

# We will run this once we are happy
#mulTree(mulTree.data = cff_mul, 
#        formula = log10(CFF) ~ log10(body_mass_us)*forage_light_level 
#                                  + mode_of_life
#                                  + habitat 
#        , parameters = parameters, 
#        chains = 2,
#        priors = prior,
#        convergence = 1.1, 
#        ESS = 1000, 
#        verbose = TRUE,
#        output = "mulTree_models", 
#        warn = FALSE, 
#        ask = TRUE)


full_model_1 <- MCMCglmm(log10(CFF) ~ log10(body_mass_us)*forage_light_level 
                                  + mode_of_life
                                  + habitat
                                  + Method,
                         random = ~ animal + Species_phylo,
                         rcov=~ units, 
                         data = cff_clean, 
                         pedigree = cff_tree[[1]], 
                         prior = prior,
                         family=c("gaussian"), 
                         nitt = nitt, thin = thin, burnin = burnin,
                         verbose= FALSE)

summary(full_model_1)

```

#Read in the multiple runs

This needs to be fixed for when we run multiple trees

```{r glm, include=FALSE}

#Main_models <-   read.mulTree(mulTree.chain = "mulTree_models", 
#                              convergence = FALSE, 
#                              model = FALSE,
#                              extract = NULL)


#inter_main_p <-  hdr(Main_models$`(Intercept)`)
#mass_log10_main_p <-  hdr(Main_models$`log10(Bodymass.g.)`)
#lit_low_main_p <-  hdr(Main_models$Light_Levellow)
#ballistic_main_p <-  hdr(Main_models$mode_of_lifepred_ballistic)
#pur_main_p <-  hdr(Main_models$mode_of_lifepred_pursuit)
#pel_main_p <-  hdr(Main_models$habitatpelagic)
#dem_main_p <-  hdr(Main_models$habitatdemersal)
#vol_main_p <-  hdr(Main_models$habitatvolant)
#body_lit_inter_main_p <-  hdr(Main_models$`log10(Bodymass.g.):Light_Levellow`)

#phylo_main_p <-  hdr(Main_models$phylogenetic.variance)

#H_main <- Main_models$phylogenetic.variance/c(Main_models$phylogenetic.variance + Main_models$residual.variance)
  
#unit_main_p <-  hdr(Main_models$residual.variance)

```


plot habitat

```{r plot habitat, include=FALSE}

plot(log10(cff_clean[, "CFF"]) ~ jitter(as.numeric(cff_clean[, "habitat"])),
       col = "white",
       pch= 16,
     bty = "n",
     xlim = c(0.5,4.5),
      xaxt = "n",
     ylab = "Log10 CFF",
     )

axis(1, at=1:4, labels= c("terrestrial", 
                          "nektonic", 
                          "demersal", 
                          "volant"))

points(log10(cff_clean[cff_clean$forage_light_level == "low", "CFF"]) ~ jitter(as.numeric(cff_clean[cff_clean$forage_light_level == "low", "habitat"]), 0.5),
       col = "grey70",
       pch= 16)

points(log10(cff_clean[cff_clean$forage_light_level == "high", "CFF"]) ~ jitter(as.numeric(cff_clean[cff_clean$forage_light_level == "high", "habitat"]), 0.5),
       col = rgb(247/250,224/250,89/250),
       pch= 16)

```


plot foraging type

```{r plot forage, include=FALSE}

plot(log10(cff_clean[, "CFF"]) ~ jitter(as.numeric(cff_clean[, "mode_of_life"])),
       col = "white",
       pch= 16,
     bty = "n",
     xlim = c(0.5,3.5),
      xaxt = "n",
     ylab = "Log10 CFF",
     )

axis(1, at=1:3, labels= c("forage", 
                          "ballistic", 
                          "pursuit"))

points(log10(cff_clean[cff_clean$forage_light_level == "low" & cff_clean$habitat != "volant", "CFF"]) ~ jitter(as.numeric(cff_clean[cff_clean$forage_light_level == "low" & cff_clean$habitat != "volant", "mode_of_life"]), 0.5),
       col = "grey70",
       pch= 16)

points(log10(cff_clean[cff_clean$forage_light_level == "high" & cff_clean$habitat != "volant", "CFF"]) ~ jitter(as.numeric(cff_clean[cff_clean$forage_light_level == "high" & cff_clean$habitat != "volant", "mode_of_life"]), 0.5),
       col = rgb(247/250,224/250,89/250),
       pch= 16)



points(log10(cff_clean[cff_clean$forage_light_level == "low" & cff_clean$habitat == "volant", "CFF"]) ~ jitter(as.numeric(cff_clean[cff_clean$forage_light_level == "low" & cff_clean$habitat == "volant", "mode_of_life"]), 0.5),
       col = "grey70",
       pch= 17)

points(log10(cff_clean[cff_clean$forage_light_level == "high" & cff_clean$habitat == "volant", "CFF"]) ~ jitter(as.numeric(cff_clean[cff_clean$forage_light_level == "high" & cff_clean$habitat == "volant", "mode_of_life"]), 0.2),
       col = rgb(247/250,224/250,89/250),
       pch= 17)



```





##terrestrial

#MulTree
Set the data into a Multree object.

```{r Multree data, include=FALSE}


cff_t <- cff_clean[cff_clean$habitat %in% c("volant", 
                                        "terrestrial"
                                        ),]

cff_t$habitat <- factor(cff_t$habitat, 
                                  levels = c("terrestrial",   
                                             "volant"
                                             ))

cff_mul_terr <- as.mulTree(cff_t, 
                      cff_tree, 
                      taxa = "animal", 
                      rand.terms = ~animal,
                      clean.data = FALSE)


```



Run the terrestrial multree model

```{r multre terr, include=FALSE}


#mulTree(mulTree.data = cff_mul_terr, 
#        formula = lgo10(CFF) ~ log10(body_mass_us)*forage_light_level 
#                                  + mode_of_life
#                                  + habitat,
#        parameters = parameters, 
#        chains = 2,
#        priors = prior,
#        convergence = 1.1, 
#        ESS = 1000, 
#        verbose = TRUE,
#        output = "cff_terr_models", 
#        warn = FALSE, 
#        ask = TRUE)


full_ter_1 <- MCMCglmm(log10(CFF) ~ log10(body_mass_us)*forage_light_level 
                                  + mode_of_life
                                  + habitat
                                  + Method,
                         random = ~ animal + Species_phylo,
                         rcov=~ units, 
                         data = cff_t, 
                         pedigree = cff_tree[[1]], 
                         prior = prior,
                         family=c("gaussian"), 
                        nitt = nitt, thin = thin, burnin = burnin,
                         verbose= FALSE)
summary(full_ter_1)

```

Read in the multiple terrestrial models 

```{r glm, include=FALSE}

#ter_models <-   read.mulTree(mulTree.chain = "cff_terr_models", 
#                              convergence = FALSE, 
#                              model = FALSE,
#                              extract = NULL)


#inter_ter_p <-  hdr(ter_models$`(Intercept)`)
#mass_log10_ter_p <-  hdr(ter_models$`log10(body_mass_us)`)
#lit_low_ter_p <-  hdr(ter_models$forage_light_levellow)
#ballistic_ter_p <-  hdr(ter_models$mode_of_lifepred_ballistic)
#pur_ter_p <-  hdr(ter_models$mode_of_lifepred_pursuit)
#vol_ter_p <-  hdr(ter_models$habitatvolant)
#body_lit_inter_ter_p <- hdr(ter_models$`log10(body_mass_us):forage_light_levellow`)

#phylo_ter_p <-  hdr(ter_models$phylogenetic.variance)

#H_ter <- ter_models$phylogenetic.variance/c(ter_models$phylogenetic.variance + #ter_models$residual.variance)
  
#unit_ter_p <-  hdr(ter_models$residual.variance)


```



Plot volant versus non volant

```{r volatn jitter, include=FALSE}

plot(log10(cff_t[, "CFF"]) ~ jitter(as.numeric(cff_t[, "habitat"])),
       col = "white",
       pch= 16,
     bty = "n",
     xlim = c(0.5,2.5),
      xaxt = "n",
     ylab = "Log10 CFF",
     )

axis(1, at=1:2, labels= c("terrestrial", 
                          "volant"))

points(log10(cff_t[cff_t$forage_light_level == "low", "CFF"]) ~ jitter(as.numeric(cff_t[cff_t$forage_light_level == "low", "habitat"]), 0.5),
       col = "grey70",
       pch= 16)

points(log10(cff_t[cff_t$forage_light_level == "high", "CFF"]) ~ jitter(as.numeric(cff_t[cff_t$forage_light_level == "high", "habitat"]), 0.5),
       col = rgb(247/250,224/250,89/250),
       pch= 16)

```


##Marine analysis


```{r marine data, include=FALSE}

cff_m <- cff_clean[cff_clean$habitat %in% c("pelagic", 
                                        "demersal"
                                        ),]

cff_m$habitat <- factor(cff_m$habitat, 
                                  levels = c("demersal",
                                             "pelagic" 
                                        ))


cff_mul_mar <- as.mulTree(cff_m, 
                      cff_tree, 
                      taxa = "animal", 
                      rand.terms = ~animal,
                      clean.data = FALSE)


```




```{r marine model, include=FALSE}


#mulTree(mulTree.data = cff_mul_mar, 
#        formula = CFF ~ log10(body_mass_us)*forage_light_level 
#                                  + mode_of_life
#                                  + habitat,
#        parameters = parameters, 
#        chains = 2,
#        priors = prior,
#        convergence = 1.1, 
#        ESS = 1000, 
#        verbose = TRUE,
#        output = "cff_mar_models", 
#        warn = FALSE, 
#        ask = TRUE)


full_mar_1 <- MCMCglmm(log10(CFF) ~ log10(body_mass_us)*forage_light_level 
                                  + mode_of_life
                                  + habitat
                                  + Method
                       ,
                         random = ~ animal + Species_phylo,
                         rcov=~ units, 
                         data = cff_m, 
                         pedigree = cff_tree[[1]], 
                         prior = prior,
                         family=c("gaussian"), 
                         nitt = nitt, thin = thin, burnin = burnin,
                         verbose= FALSE)
summary(full_mar_1)

```



Read in the multiple marine models 

```{r marine read, include=FALSE}

mar_models <-   read.mulTree(mulTree.chain = "cff_mar_models", 
                              convergence = FALSE, 
                              model = FALSE,
                              extract = NULL)


inter_mar_p <-  hdr(mar_models$`(Intercept)`)
mass_log10_mar_p <-  hdr(mar_models$`log10(Bodymass.g.)`)
lit_low_mar_p <-  hdr(mar_models$Light_Levellow)
ballistic_mar_p <-  hdr(mar_models$mode_of_lifepred_ballistic)
#scav_mar_p <-  hdr(mar_models$mode_of_lifescavenger)
pur_mar_p <-  hdr(mar_models$mode_of_lifepred_pursuit)
body_lit_inter_mar_p <-  hdr(mar_models$`log10(Bodymass.g.):Light_Levellow`)

phylo_mar_p <-  hdr(mar_models$phylogenetic.variance)

H_mar <- mar_models$phylogenetic.variance/c(mar_models$phylogenetic.variance + mar_models$residual.variance)
  
unit_mar_p <-  hdr(mar_models$residual.variance)


```



plot marine
```{r glm, include=FALSE}



plot(cff_m[, "Max_CFF_Hz"] ~ log10(cff_m[, "Bodymass.g."]),
       col = "white",            
        cex = 1.3,
       pch= 16,
     bty = "n")


points(cff_m[cff_m$Light_Level == "high", "Max_CFF_Hz"] ~ log10(cff_m[cff_m$Light_Level == "high", "Bodymass.g."]),
       col = rgb(247/250,224/250,89/250),
       pch= 16)

points(cff_m[cff_m$Light_Level == "low", "Max_CFF_Hz"] ~ log10(cff_m[cff_m$Light_Level == "low", "Bodymass.g."]),
       col = "grey70",
       pch= 16)


points(cff_m[cff_m$mode_of_life == "pred_ballistic", "Max_CFF_Hz"] ~ log10(cff_m[cff_m$mode_of_life == "pred_ballistic", "Bodymass.g."]),
       col = "white",
       cex = 0.7,
       pch= 16)




points(cff_m[cff_m$mode_of_life == "pred_pursuit" & cff_m$Light_Level == "high", "Max_CFF_Hz"] ~ log10(cff_m[cff_m$mode_of_life == "pred_pursuit" & cff_m$Light_Level == "high", "Bodymass.g."]),
       col = "white",
       cex = 1.3,
       pch= 16)

points(cff_m[cff_m$mode_of_life == "pred_pursuit" & cff_m$Light_Level == "low", "Max_CFF_Hz"] ~ log10(cff_m[cff_m$mode_of_life == "pred_pursuit" & cff_m$Light_Level == "low", "Bodymass.g."]),
       col = "white",
       cex = 1.3,
       pch= 17)




points(cff_m[cff_m$mode_of_life == "pred_pursuit" & cff_m$Light_Level == "high", "Max_CFF_Hz"] ~ log10(cff_m[cff_m$mode_of_life == "pred_pursuit" & cff_m$Light_Level == "high", "Bodymass.g."]),
       col = rgb(247/250,224/250,89/250),
       cex = 1,
       pch= 17)

points(cff_m[cff_m$mode_of_life == "pred_pursuit" & cff_m$Light_Level == "low", "Max_CFF_Hz"] ~ log10(cff_m[cff_m$mode_of_life == "pred_pursuit" & cff_m$Light_Level == "low", "Bodymass.g."]),
       col = "grey70",
       cex = 1,
       pch= 17)



##high light
abline(summary(full_mar_1)$solutions[1], summary(full_mar_1)$solutions[2],
       col = rgb(247/250,224/250,89/250),
       lty = 4,
       lwd = 3)

##low light
abline(summary(full_mar_1)$solutions[1] + summary(full_mar_1)$solutions[3], 
       summary(full_mar_1)$solutions[2] + summary(full_mar_1)$solutions[8],
       col = "grey70",
       lty = 4,
       lwd = 3)


```




Plot ballistic versus non volant

```{r volatn jitter, include=FALSE}


plot(log10(cff_m[, "CFF"]) ~ jitter(as.numeric(cff_m[, "mode_of_life"])),
       col = "white",
       pch= 16,
     bty = "n",
     xlim = c(0.5,3.5),
      xaxt = "n",
     ylab = "Log10 CFF",
     )

axis(1, at=1:3, labels= c("forage", 
                          "ballistic", 
                          "pursuit"))

points(log10(cff_m[cff_m$forage_light_level == "low", "CFF"]) ~ jitter(as.numeric(cff_m[cff_m$forage_light_level == "low" & cff_m$habitat != "volant", "mode_of_life"]), 0.5),
       col = "grey70",
       pch= 16)

points(log10(cff_m[cff_m$forage_light_level == "high", "CFF"]) ~ jitter(as.numeric(cff_m[cff_m$forage_light_level == "high", "mode_of_life"]), 0.5),
       col = rgb(247/250,224/250,89/250),
       pch= 16)


```


```






#plot phylo
We can also plot the phylogeny

```{r phylo plot}

#First make a named vector of the mean trait value.

plot_CFF <- vector()

for(i in 1:length(cff_clean$Scientific_name)){
  
  plot_CFF[i] <-  cff_clean[ cff_clean$Scientific_name ==  sort(cff_clean$Scientific_name)[i] , "Max_CFF_Hz"]

}

names(plot_CFF) <- sort(cff_tree[[1]]$tip.label)


#pdf("phylobar_ld50.pdf")
plotTree.wBars(tree = cff_tree[[1]], 
               x = plot_CFF, 
               scale= 1, 
               type="fan", 
               cex = 0.2, 
               tip.labels = T, 
               fsize=0.37, 
               width = 1)
#dev.off()

```
